---
title: Covid Cases
---

#### Outcomes

- Download and load public data sets
- Read and interpret metadata
- Evaluate general claims using data, in the presence of missing data
- Create and interpret interaction plots


#### Instructions

- Answer the following questions, and show all your R code.
- Upload your submission to Canvas in nicely formatted HTML generated from Rstudio.

Download the [COVID-19 Case Surveillance Public Use Data](https://data.cdc.gov/Case-Surveillance/COVID-19-Case-Surveillance-Public-Use-Data/vbim-akqf) from the US Center for Disease Control (CDC).
Click on the "Export" button in the top right corner next to download the actual data set.

_Hint: It might be useful to create a new numeric column for severity based on the values of the `hosp, icu, death` columns.
If you do this, make sure you explain the meaning and rationale behind your choices._


# Questions

## 1 - Reading the Documentation

_a) Describe this data set at a high level.
What does each row represent?_

This data set includes COVID-19 cases reported in US with date, medical condition data, ICU data, number of deaths, sex, age groups, race and ethnicity, etc. Each row represents each case reported.

_b) How many rows and columns do you expect to see based on the website?
Load the data into R.
Is this reflected in the data?_

```{r}
covid = read.csv("C:/Users/kanni/Downloads/COVID-19_Case_Surveillance_Public_Use_Data.csv")
dim(covid)
```

There are 4481062 rows and 11 columns in the data.


## 2 - Missing Data
 
_There are few `NA` values in this data, but much of the data is missing, unknown, or blank.
Which columns have the highest proportion of missing data?
Why might these be missing?_

```{r}
missing_count = colSums(is.na(covid) | covid == "Missing" | covid == "Unknown")
missing_count
```

Counting all `NA` values which includes missing, unknown, the column with highest counts is icu_yn with 3,932,347 `NA` values. These data might be missing as patients are still under care for COVID, but haven't yet to be in ICU.


## 3 - Age Risk

_The [CDC states](https://www.cdc.gov/coronavirus/2019-ncov/need-extra-precautions/older-adults.html) "Among adults, the risk for severe illness from COVID-19 increases with age, with older adults at highest risk."
Does the data support this claim?_

_Hint 1: It might be useful to think about these COVID illness rates by age after adjusting     for the relative sizes of each age group in the country.
You'll need to bring in another data source to do this._


```{r}
severity = rep(0, times = nrow(covid))
severity[covid$hosp_yn == "Yes"] = 1
severity[covid$icu_yn == "Yes"] = 2
severity[covid$death_yn == "Yes"] = 3
covid$severity = severity

table(covid$age_group)
Age_Severe_Plot = table(covid[, c("age_group", "severity")])
plot(table(covid$age_group),main = "Confirmed Cases by Age Groups", xlab = "Age Groups", ylab ="Confirmed Cases")

plot(Age_Severe_Plot,main = "Severe Cases by Age Groups",xlab="Age Groups",ylab="Severity Level")
```

For severity, I select `hosp_yn` as first level of severe illness, `icu_yn` as second level, `death_yn` as third level, and the rest as no severe illness. The table shows the number of cases at first, second, third level of severe illness increase with the ages groups. Also, cases with no severe illness decrease with the age groups.


## 4 - Medical Condition Risk

_The [CDC states](https://www.cdc.gov/coronavirus/2019-ncov/need-extra-precautions/people-with-medical-conditions.html) "People of any age with certain underlying medical conditions are at increased risk for severe illness from COVID-19"
Does the data support this claim?_

```{r}
table(covid$medcond_yn)
medcond_Severe_Plot = table(covid[, c("medcond_yn", "severity")])
plot(table(covid$medcond_yn),main = "Confirmed Cases with Pre-Medical Conditions", xlab = "Pre-Medical Conditions", ylab ="Confirmed Cases")

plot(medcond_Severe_Plot,main = "Severe Cases with Pre-Medical Conditions",xlab="Pre-Medical Conditions",ylab="Severity Level")
```

The table shows there are more cases at 1st-3rd level of severe illness among patients with pre-medical conditions. Also, there are more cases at no risk of severe illness among patients with no pre-medical conditions. This might not be enough to conclude patients with pre-medical conditions would be at a higher risk of severe illness. Besides, for `NA` values, it is still a relatively low number of cases at higher risk of severe illness.


## 5 - Interaction Effects

_In statistics, an __interaction effect__ is when two or more variables have a non additive effect in the value of some response.
In the questions above, we explored the severity of COVID-19 (response) given variables age and medical condition.
Do age and medical condition have an interaction effect in the severity of COVID-19?_

```{r}
class(severity)
with(covid, interaction.plot(age_group, medcond_yn, severity))
```

The graph indicates this might be the case that age groups and pre-medical conditions do have an interactive effect with the risk of severe illness.


## 6 - Your Question

_Come up with your own question about COVID-19, and answer it using this data set.
You may want to explore other demographic columns: sex, and race / ethnicity.
You could also look at how the number of cases have changed over time using the report dates._

```{r}
covid$medical = as.numeric(covid$medcond_yn == "Yes")
with(covid, interaction.plot(medical, sex, severity))
```

This graph includes the risk of severe illness as a response with given variables as pre-medical conditions and sex. The graph shows this might be the case that pre-medical conditions and sex do have an interactive effect with the risk of severe illness. In the case of missing, unknown, and others, it is inconclusive to interpret there is a shift.




