---
title: "IntroRmarkdown"
Author: "Kannika Pho"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Outcomes
+ Create HTML report using Rstudio
+ Embed plots in HTML
+ Manipulate named vectors

1. Average Annual Precipitation by cities in US is measured in [inches/milimeters with ranks](https://www.currentresults.com/Weather/US/average-annual-precipitation-by-city.php)

2. Verify Sacramento is in the data.
```{r}
x = (precip)
y = (names(x))
"Sacramento" %in% y
```
  Yes, Sacramento is in the data using %in%

3. Which 5 cities receive the most rain? The least rain?
```{r}
z = sort(x)
z
tail(z)
head(z)
```
  5 cities receive the most rain are Mobile, Miami, San Juan, New Orleans, Juneau, and Jacksonville. 5 cities receive the least rain are Phoenix, Reno, Albuquerque, El Paso, Boise, and Denver.
```{r}
r = z[70]/z[1]
d = x - mean(x)
d = abs(d)
sort(d)
hist(x)
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
