---
title: Midterm
author: "Kannika Pho"
date: 1 November 2020
---

```{r echo=FALSE, results='hide',message=FALSE}
library(ggplot2)

# Just ignore this code :)
.score = new.env()
.score$PASS = 0
.score$FAIL = 0

#' Check that your answer is correct.
#'
#' @param expr expression that should evaluate to TRUE if the answer is correct.
check = function(expr)
{
  status = tryCatch(expr, error = identity)
  if(isTRUE(status)){
    .score$PASS = .score$PASS + 1
    cat("PASS")
  } else {
    .score$FAIL = .score$FAIL + 1
    cat("FAIL")
  }
}

#' Check that every element of x and y are "close enough".
#'
#' @param x numeric vector
#' @param y numeric vector
close_enough = function(x, y, epsilon = 1e-5)
{
  max(abs(x - y)) < epsilon
}
```


#### Instructions

- Leave all the calls to `check()` where they are, and use these to verify that you got the right answer.
- Answer the following questions, and show all your R code.
- Upload your submission to Canvas in nicely formatted HTML generated from Rstudio.


## 1

(2 pts)

Use `seq` to create a vector `x` containing the sequence -2.0, -1.9, ..., 4.9, 5.0.

```{r}
x = seq(-2.0, 5.0, by=0.1 )

check(x[1] == -2)
check(length(x) == 71)
```

## 2

(4 pts)

The function `dunif` in R evaluates the probability density function (PDF) for the uniform distribution on the interval from ‘min’ to ‘max’.

- Call `dunif` on `x` to create and plot three lines with different colors on the same graph.
- Add titles and axis labels as appropriate.

Your plot should include one line for the PDF of a Uniform(0, 1) distribution, one line for the PDF of a Uniform(-1, 2) distribution, and one line for the PDF of a Uniform(0.5, 2.5) distribution.
Hint: You can compute the PDF of a Uniform(a, b) distribution with `dunif(..., min = a, max = b)`.

```{r}
d1 = dunif(x, min = 0, max = 1)
d2 = dunif(x, min = -1, max = 2)
d3 = dunif(x, min = 0.5, max = 2.5)
d = data.frame(x,d1,d2,d3)

g = ggplot(data = d, mapping = aes(x)) + geom_line(aes(y = d1, colour= "Uniform(0,1)")) + geom_line(aes(y = d2, colour= "Uniform(-1,2)")) + geom_line(aes(y = d3, colour= "Uniform(0.5,2.5)")) + ylab("Probability Density Function (PDF)")  + ggtitle("Probability Density Function Evaluation for Uniform Distribution")

print(g)
```


## 3

(5 pts)

Define and implement a function `dunif2` that accepts the arguments `x`, `min`, `max`, and behaves exactly like `dunif`.
Don't do `dunif2 = dunif`, that's cheating :)
Don't worry about the `log` argument to `dunif`.

```{r}
dunif2 = function(x = seq(-2.0, 5.0, by=0.1 ), min = 0, max = 1)
{
  for(i in seq_along(x))
  {
    if ( x[i] <= min | x[i] >= max)
    {
      x[i] = NaN
    } else
      x[i] = 1/(max-min)
  }
}

x3 = seq(from = -10, to = 10)
check(close_enough(
  dunif(x3), dunif2(x3)
))
check(close_enough(
  dunif(x3, min = -0.5, max = 2.5), 
  dunif2(x3, min = -0.5, max = 2.5)
))
```

## 4

(3 pts)

Define `mtcars2` by selecting the rows of `mtcars` that have `mpg` higher than the median value of `mpg`, and have `cyl` equal to 4.

```{r}
data(mtcars)
mtcars2 = mtcars[median(mtcars$mpg) < mtcars$mpg & 4 == mtcars$cyl, ]

check(nrow(mtcars2) == 11)
```


## 5

(2 pts)

In the following broken code, the goal was to produce a scatter plot of the `mtcars` data frame with `disp` on the x axis, and `mpg` on the y axis, and color determined by `gear`.
Fix the code so it makes the correct plot.

```{r}
g = ggplot(mtcars, mapping = aes(x = disp, y = mpg)) +
  geom_point(aes(color = as.factor(gear)))

print(g)
```

## 6

(4 pts)

Suppose you want to know the length of several objects in a list, so you write the following code:

```{r}
l = list(letters, LETTERS, 1:10, rnorm(5))

l_len = c(length(l[[1]]), length(l[[2]]), length(l[[3]]), length(l[[4]]))
```

What is wrong with the above approach to calculating the length of each element?

Calculating the length of each element above requires typing each element in the list, long and more chances getting bugs. lapply uses less code which means less opportunities for bugs. lapply is fast, idiomatic, and easy to make faster through parallel programming.

Define `l_len2` by computing the lengths of all elements in `l` in an idiomatic way.

```{r}
l_len2 = do.call(rbind, lapply(l, length))

check(close_enough(l_len, l_len2))
```


## Auto Score

Your score below doesn't include parts of the assignment that aren't automatically graded, for example, plots, short answers, and data interpretation.

```{r echo=FALSE}
sprintf("Auto score:  %i / %i total points", .score$PASS, .score$PASS + .score$FAIL)
```